---
title: "rix: reproducible development environments for R programmers"
author: "Bruno Rodrigues"
format:
  revealjs:
    embed-resources: true
    slide-number: true
    theme: solarized
    incremental: true
execute:
  echo: true
---

## What is R?

- GNU R is a multi-paradigm programming language (first released 1993)
- FLOSS implementation of the S language developed at Bell Labs
- Statistics, machine learning and data science
- Great for visualisation tasks as well
- Built-in objects: data frames, matrices, formulas, models...

## Hello World!

```{r, include = F}
targets::tar_read(life_cycle_savings)

fm1 <- lm(sr ~ pop15 + pop75 + dpi + ddpi, data = LifeCycleSavings)

summary(fm1)
```

```{r, eval = F}
life_cycle_savings <- read.csv("life_cycle_savings.csv")

fm1 <- lm(sr ~ pop15 + pop75 + dpi + ddpi, data = LifeCycleSavings)

summary(fm1)
```

## CRAN and Bioconductor

- CRAN is the repository of R packages to extend the language
- As of writing, 20331 packages available
- Biocondcutor: repository with a focus on Bioinformatics: 2266 more packages
- All available through `nixpkgs` in the `rPackages` set!

## Per-project environments

not really an R thing
mostly focused on per-project library of packages

## rix

per-project reproducible environments
rix()
with_nix()

## To know more

https://b-rodrigues.github.io/rix/
